---
title: "CPP 526: Final Project"
output: 
  flexdashboard::flex_dashboard:
    theme: spacelab
    source: embed
    smart: false
runtime: shiny
---

```{r include = FALSE}

# LOAD PACKAGES

library( DT)
library( ggmap )
library( shiny )
library( knitr )
library( pander )
library( leaflet )
library( viridis )
library( rsconnect )
library( tidyverse )
library( flexdashboard )
library(ggplot2)



# READ IN DATA

url <- paste0("https://github.com/DS4PS/Data-",
              "Science-Class/blob/master/DATA",
              "/TempeTrafficAccidents.rds?raw=true")

dat <- readRDS( gzcon( url( url ) ) )



# DATA PREPROCESSING I: INJURIES & FATALITIES

dat        <- na.omit(dat)                                       # Remove NA
dat$fatal  <- dat$Totalfatalities > 0                            # 1+ fatalities
dat$inj    <- dat$Totalinjuries > 0 & dat$Totalfatalities == 0   # 1+ injuries, 0 fatal
dat$nohurt <- dat$Totalfatalities + dat$Totalinjuries == 0       # Harmless

date.vec   <- strptime(dat$DateTime, format = "%m/%d/%y %H:%M")  # Datetime variables
dat$hour   <- format(date.vec, format = "%H") %>% as.numeric()
dat$month  <- format(date.vec, format = "%b")
dat$day    <- format(date.vec, format = "%a")
dat$day365 <- format(date.vec, format = "%j")
dat$week   <- format(date.vec, format = "%V")



# DATA PREPROCESSING II: NAMED INTERVALS OF TIME

dat <- dat %>% 
  mutate(time.of.day = case_when(hour >= 6 & hour <= 9 ~ "Morning Commute", 
                                 hour >= 16 & hour <= 19 ~ "Evening Commute", 
                                 hour >= 14 & hour <= 15 ~ "School Pickup", 
                                 hour >= 9 & hour <= 13 ~ "Work", 
                                 hour >= 20 & hour <= 23 ~ "Night", 
                                 hour <= 5 & hour >= 0 ~ "Midnight to Dawn"))

dat$harm <- ifelse(test = dat$Totalinjuries > 0 | dat$Totalfatalities > 0, 
                   yes = "Harm", 
                   no = "No Harm")



# DATA PREPROCESSING III: PERMUTATIONS OF INEBRIATION

dat <- dat %>% 
  mutate(d1.substance = case_when(AlcoholUse_Drv1 == "Alcohol" & 
                                  DrugUse_Drv1 == "No Apparent Influence" ~ "Alcohol", 
                                  AlcoholUse_Drv1 == "No Apparent Influence" & 
                                  DrugUse_Drv1 == "Drugs" ~ "Drugs", 
                                  AlcoholUse_Drv1 == "Alcohol" & 
                                  DrugUse_Drv1 == "Drugs" ~ "Alcohol and Drugs", 
                                  AlcoholUse_Drv1 == "No Apparent Influence" & 
                                  DrugUse_Drv1 == "No Apparent Influence" ~ "No Apparent Influence"))

dat <- dat %>% 
  mutate(d2.substance = case_when(AlcoholUse_Drv2 == "Alcohol" & 
                                  DrugUse_Drv2 == "No Apparent Influence" ~ "Alcohol", 
                                  AlcoholUse_Drv2 == "No Apparent Influence" & 
                                  DrugUse_Drv2 == "Drugs" ~ "Drugs", 
                                  AlcoholUse_Drv2 == "Alcohol" & 
                                  DrugUse_Drv2 == "Drugs" ~ "Alcohol and Drugs", 
                                  AlcoholUse_Drv2 == "No Apparent Influence" & 
                                  DrugUse_Drv2 == "No Apparent Influence" ~ "No Apparent Influence"))



# DATA PREPROCESSING IV: AGE CATEGORIES

dat$age.cat <- case_when(dat$Age_Drv1 >= 0 & 
                         dat$Age_Drv1 <= 18 ~ "Youth", 
                         dat$Age_Drv1 >= 19 & 
                         dat$Age_Drv1 <= 25 ~ "Young Adult", 
                         dat$Age_Drv1 >= 26 & 
                         dat$Age_Drv1 <= 64 ~ "Adult", 
                         dat$Age_Drv1 >= 65 ~ "Senior")

```

Day & Time
=====================================  

Inputs {.sidebar}
-------------------------------------

```{r}

checkboxGroupInput("days", 
                   label = h3("Day of Week"), 
    choices = list("Monday"    = "Mon", 
                   "Tuesday"   = "Tue", 
                   "Wednesday" = "Wed", 
                   "Thursday"  = "Thu",
                   "Friday"    = "Fri",
                   "Saturday"  = "Sat",
                   "Sunday"    = "Sun" ),
    selected = c("Fri",
                 "Sat",
                 "Sun")) #Selected is the default that loads

sliderInput(inputId = "hour", #column variable it is based off of
            label = h3("Time of Day"), #What is h3? 
            min = 0, 
            max = 23, 
            value = c(6, 12)) #This is the default that loads


```

Outputs
-------------------------------------

### Traffic Accidents By Day and Time

```{r}

# LEAFLET MAPPING

renderLeaflet({
  
  d2 <- dat %>%
    filter(day %in% input$days, 
           hour >= input$hour[1],
           hour <= input$hour[2])
  
  d2$col.vec <- ifelse(test = d2$nohurt, 
                       yes = "gray20", 
                       no = ifelse(test = d2$inj, 
                                   yes = "steelblue", 
                                   no = "darkorange") ) 
  
  point.size <- d2$Totalinjuries + d2$Totalfatalities

  crash.details <- paste0("Time: ", d2$DateTime, "<br>",
                          "Total Fatalities: ", d2$Totalfatalities, "<br>",
                          "Total Injuries: ", d2$Totalinjuries, "<br>",
                          "Collision type: ", d2$Collisionmanner)
  
  tempe <- leaflet( ) %>% 
    addProviderTiles("CartoDB.Positron")  %>%
    setView(lng = -111.9278, 
            lat = 33.39951, 
            zoom = 13)
  
  addCircles(tempe, 
             lng = d2$Longitude, 
             lat = d2$Latitude,
             fillColor = d2$col.vec, 
             fillOpacity = 0.5, 
             stroke = FALSE, 
             radius = 50 * (1 + 0.33 * point.size),
             popup = crash.details)

})

```   

Age, Gender, & Method {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
-------------------------------------
Driver Characteristics

```{r}

sliderInput(inputId = "d1age", 
            label = h4("Driver 1 Age"), 
            min = 15, 
            max = 100, 
            value = c(18,36) )

sliderInput(inputId = "d2age", 
            label = h4("Driver 2 Age"), 
            min = 15, 
            max = 100, 
            value = c(18,36) )

selectInput(inputId = "d1gender", 
            label = h4("Driver 1 Gender"), 
            choices = c("Male",
                        "Female", 
                        "Unknown"), 
            selected = "Male")

selectInput(inputId = "d2gender", 
            label = h4("Driver 2 Gender"), 
            choices = c("Male",
                        "Female", 
                        "Unknown"), 
            selected = "Male")

radioButtons(inputId = "d1pedcy", 
             label = h4("Driver 1 Transportation"),
             choices = c("Driver", 
                         "Pedalcyclist", 
                         "Pedestrian"), 
             selected = "Driver")

radioButtons(inputId = "d2pedcy", 
             label = h4("Driver 2 Transportation"),
             choices = c("Driver", 
                         "Pedalcyclist", 
                         "Pedestrian"), 
             selected = "Driver")

```

Row 
-------------------------------------

### Number of Crashes
  
```{r}

renderValueBox({
  
  d2 <- dat %>%
    filter(Age_Drv1 >= input$d1age[1],
           Age_Drv1 <= input$d1age[2], 
           Age_Drv2 >= input$d2age[1],
           Age_Drv2 <= input$d2age[2], 
           Gender_Drv1 %in% input$d1gender, 
           Gender_Drv2 %in% input$d2gender, 
           Unittype_One %in% input$d1pedcy, 
           Unittype_Two %in% input$d2pedcy )
  
  crashes <- count(d2)
  
  valueBox(crashes, 
           icon = "fa-solid fa-car-burst",
           color = ifelse(test = crashes > 50, 
                          yes = "danger", 
                          no = "primary") )
  
})

```

### Total Injuries
  
```{r}

renderValueBox({
  
  d2 <- dat %>%
    filter(Age_Drv1 >= input$d1age[1],
           Age_Drv1 <= input$d1age[2], 
           Age_Drv2 >= input$d2age[1],
           Age_Drv2 <= input$d2age[2], 
           Gender_Drv1 %in% input$d1gender, 
           Gender_Drv2 %in% input$d2gender, 
           Unittype_One %in% input$d1pedcy, 
           Unittype_Two %in% input$d2pedcy)
  
  total.injuries <- sum(d2$Totalinjuries)
  
  valueBox(total.injuries, 
           icon = "fa-solid fa-user-injured",
           color = ifelse(test = total.injuries > 30, 
                          yes = "danger", 
                          no = "primary" ))

  })

```

### Total Fatalities
  
```{r}

renderValueBox({
  
  d2 <- dat %>%
    filter(Age_Drv1 >= input$d1age[1],
           Age_Drv1 <= input$d1age[2], 
           Age_Drv2 >= input$d2age[1],
           Age_Drv2 <= input$d2age[2], 
           Gender_Drv1 %in% input$d1gender, 
           Gender_Drv2 %in% input$d2gender, 
           Unittype_One %in% input$d1pedcy, 
           Unittype_Two %in% input$d2pedcy)

  total.fatalities <- sum(d2$Totalfatalities)
  
  valueBox(total.fatalities, 
           icon = "fa-solid fa-skull",
           color = ifelse(test = total.fatalities > 10, 
                          yes = "danger", 
                          no = "primary"))
  
})

```

### Rate of Harm
  
```{r}

renderValueBox({
  
  d2 <- dat %>%
    filter(Age_Drv1 >= input$d1age[1],
           Age_Drv1 <= input$d1age[2], 
           Age_Drv2 >= input$d2age[1],
           Age_Drv2 <= input$d2age[2], 
           Gender_Drv1 %in% input$d1gender, 
           Gender_Drv2 %in% input$d2gender, 
           Unittype_One %in% input$d1pedcy, 
           Unittype_Two %in% input$d2pedcy)
  
  rate.of.harm <- round(length(which(d2$harm == "Harm")) / count(d2), 3)
  
  valueBox(rate.of.harm, 
           icon = "fa-solid fa-percent",
           color = ifelse(test = rate.of.harm > 0.5, 
                          yes = "danger", 
                          no = "primary"))
  
})

```

Outputs
-------------------------------------

### Traffic Accidents by Driver Characteristics

```{r}

renderLeaflet({
  
  d2 <- dat %>%
    filter(Age_Drv1 >= input$d1age[1],
           Age_Drv1 <= input$d1age[2], 
           Age_Drv2 >= input$d2age[1],
           Age_Drv2 <= input$d2age[2], 
           Gender_Drv1 %in% input$d1gender, 
           Gender_Drv2 %in% input$d2gender, 
           Unittype_One %in% input$d1pedcy, 
           Unittype_Two %in% input$d2pedcy)
  
  d2$col.vec <- ifelse(test = d2$nohurt, 
                       yes = "gray20", 
                       no = ifelse(test = d2$inj, 
                                   yes = "steelblue", 
                                   no = "darkorange") )              
    
  point.size <- d2$Totalinjuries + d2$Totalfatalities

  crash.details <- paste0("Time: ", d2$DateTime, "<br>",
                          "Total Fatalities: ", d2$Totalfatalities, "<br>",
                          "Total Injuries: ", d2$Totalinjuries, "<br>",
                          "Collision type: ", d2$Collisionmanner)
  
  tempe <- leaflet() %>% 
    addProviderTiles("CartoDB.Positron")  %>%
    setView(lng = -111.9278, 
            lat = 33.39951, 
            zoom = 13)
  
  
  addCircles(tempe, 
             lng = d2$Longitude, 
             lat = d2$Latitude,
             fillColor = d2$col.vec, 
             fillOpacity = 0.5, 
             stroke = FALSE, 
             radius = 50 * (1 + 0.33 * point.size),
             popup = crash.details)


})

```   

Comparisons {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
-------------------------------------
Driver Characteristics

```{r}

sliderInput(inputId = "driver.1.age", 
            label = h4("Driver 1 Age"), 
            min = 15, 
            max = 100, 
            value = c(18, 36))

sliderInput(inputId = "driver.2.age", 
            label = h4("Driver 2 Age"), 
            min = 15, 
            max = 100, 
            value = c(18, 36))

selectInput(inputId = "driver.1.gender", 
            label = h4("Driver 1 Gender"), 
            choices = c("Male",
                        "Female", 
                        "Unknown"), 
            selected = "Male")

selectInput(inputId = "driver.2.gender", 
            label = h4("Driver 2 Gender"), 
            choices = c("Male",
                        "Female",
                        "Unknown"), 
            selected = "Male")

radioButtons(inputId = "driver.1.pedcy", 
             label = h4("Driver 1 Transportation"),
             choices = c("Driver",
                         "Pedalcyclist",
                         "Pedestrian"), 
             selected = "Driver")

radioButtons(inputId = "driver.2.pedcy", 
             label = h4("Driver 2 Transportation"),
             choices = c("Driver",
                         "Pedalcyclist",
                         "Pedestrian"), 
             selected = "Driver")

```

Row 
-------------------------------------

### Number of Crashes
  
```{r}

renderValueBox({
  
  d2 <- dat %>%
    filter(Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
           Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
           Gender_Drv1 %in% input$d1gender, 
           Gender_Drv2 %in% input$d2gender, 
           Unittype_One %in% input$d1pedcy, 
           Unittype_Two %in% input$d2pedcy )
  
  crashes <- count(d2)
  
  valueBox(crashes, 
           icon = "fa-solid fa-car-burst",
           color = ifelse( crashes > 50, "danger", "primary") )
  
})

```

### Total Injuries
  
```{r}

renderValueBox({
  
  d2 <- dat %>%
    filter(Age_Drv1 >= input$d1age[1] & Age_Drv1 <= input$d1age[2], 
           Age_Drv2 >= input$d2age[1] & Age_Drv2 <= input$d2age[2], 
           Gender_Drv1 %in% input$d1gender, 
           Gender_Drv2 %in% input$d2gender, 
           Unittype_One %in% input$d1pedcy, 
           Unittype_Two %in% input$d2pedcy)
  
  total.injuries <- sum(d2$Totalinjuries)
  
  valueBox(total.injuries, 
           icon = "fa-solid fa-user-injured",
           color = ifelse(total.injuries > 30, "danger", "primary"))
  
})

```

### Total Fatalities
  
```{r}

renderValueBox({
  
  d2 <- dat %>%
    filter(Age_Drv1 >= input$d1age[1],
           Age_Drv1 <= input$d1age[2], 
           Age_Drv2 >= input$d2age[1],
           Age_Drv2 <= input$d2age[2], 
           Gender_Drv1 %in% input$d1gender, 
           Gender_Drv2 %in% input$d2gender, 
           Unittype_One %in% input$d1pedcy, 
           Unittype_Two %in% input$d2pedcy)

  total.fatalities <- sum(d2$Totalfatalities)
  
  valueBox(total.fatalities, 
           icon = "fa-solid fa-skull",
           color = ifelse(test = total.fatalities > 10, 
                          yes = "danger", 
                          no = "primary"))
  
})

```

### Rate of Harm
  
```{r}

renderValueBox({
  
  d2 <- dat %>%
    filter(Age_Drv1 >= input$d1age[1],
           Age_Drv1 <= input$d1age[2], 
           Age_Drv2 >= input$d2age[1],
           Age_Drv2 <= input$d2age[2], 
           Gender_Drv1 %in% input$d1gender, 
           Gender_Drv2 %in% input$d2gender, 
           Unittype_One %in% input$d1pedcy, 
           Unittype_Two %in% input$d2pedcy)
  
  rate.of.harm <- round(length(which(d2$harm == "Harm")) / count(d2), 3)
  
  valueBox(rate.of.harm, 
           icon = "fa-percent",
           color = ifelse(test = rate.of.harm > 0.5, 
                          yes = "danger", 
                          no = "primary"))
  
})

```

Column
-------------------------------------

### Driver 1

```{r}

renderLeaflet({
  
  d10 <- dat %>%
    filter(Age_Drv1 >= input$driver.1.age[1],
           Age_Drv1 <= input$driver.1.age[2], 
           Gender_Drv1 %in% input$driver.1.gender, 
           Unittype_One %in% input$driver.1.pedcy )
  
  d10$col.vec <- ifelse(test = d10$nohurt, 
                        yes = "gray20", 
                        no = ifelse(test = d10$inj, 
                                    yes = "steelblue", 
                                    no = "darkorange"))              
    
  point.size <- d10$Totalinjuries + d10$Totalfatalities

  crash.details <- paste0("Time: ", d10$DateTime, "<br>",
                          "Total Fatalities: ", d10$Totalfatalities, "<br>",
                          "Total Injuries: ", d10$Totalinjuries, "<br>",
                          "Collision type: ", d10$Collisionmanner)
  
  tempe <- leaflet() %>% 
    addProviderTiles("CartoDB.Positron") %>%
    setView(lng = -111.9278, 
            lat = 33.39951, 
            zoom = 13)
  
  addCircles(tempe, 
             lng = d10$Longitude, 
             lat = d10$Latitude,
             fillColor = d10$col.vec, 
             fillOpacity = 0.5, 
             stroke = FALSE, 
             radius = 50 * (1 + 0.33 * point.size),
             popup = crash.details)
  
  


})
```   

### Driver 2

```{r}

renderLeaflet({
  
  d11 <- dat %>%
    filter(Age_Drv2 >= input$driver.2.age[1],
           Age_Drv2 <= input$driver.2.age[2], 
           Gender_Drv2 %in% input$driver.2.gender, 
           Unittype_Two %in% input$driver.2.pedcy)
  
  d11$col.vec <- ifelse(test = d11$nohurt, 
                        yes = "gray20", 
                        no = ifelse(test = d11$inj, 
                                    yes = "steelblue", 
                                    no = "darkorange"))
  
  point.size2 <- d11$Totalinjuries + d11$Totalfatalities

  crash.details2 <- paste0("Time: ", d11$DateTime, "<br>",
                           "Total Fatalities: ", d11$Totalfatalities, "<br>",
                           "Total Injuries: ", d11$Totalinjuries, "<br>",
                           "Collision type: ", d11$Collisionmanner)
  
  tempe2 <- leaflet() %>% 
    addProviderTiles("CartoDB.Positron") %>%
    setView(lng = -111.9278, 
            lat = 33.39951, 
            zoom = 13)
  
  addCircles(tempe2, 
             lng = d11$Longitude, 
             lat = d11$Latitude,
             fillColor = d11$col.vec, 
             fillOpacity = 0.5, 
             stroke = FALSE, 
             radius = 50 * (1 + 0.33 * point.size2),
             popup = crash.details2 )

})

```   

Year & Harm
=====================================  

Inputs {.sidebar}
-------------------------------------
```{r}
checkboxGroupInput("Year", 
                   label = h3("Year"), 
    choices = list(2012, 
                   2013, 
                   2014, 
                   2015,
                   2016,
                   2017,
                   2018),
    selected = c(2012, 2013, 2014, 2015, 2016, 2017, 2018))

checkboxGroupInput(inputId = ("level_of_harm"),
            label = h3("Type of Accident"), 
            choices = list("Harmless" = "nohurt", 
                           "Injurious (No Fatalities)" = "inj",
                           "Fatal" = "fatal"),
            selected = c("nohurt", "inj", "fatal"))
```


Outputs
-------------------------------------

### Number of Accidents by Year & Harm

```{r}

renderTable({
  req(input$Year)
  
filtered_data <- dat %>% 
  filter(Year %in% input$Year) 
    
aggregated_data <- filtered_data %>%
  group_by(Year) %>%
  summarize(
    "Fatal Accidents" = sum(fatal, na.rm = TRUE),
    "Injurious Accidents" = sum(inj, na.rm = TRUE),
    "Harmless Accidents" = sum(nohurt, na.rm = TRUE),
    "Total Accidents" = sum(fatal, inj, nohurt, na.rm = TRUE),
    "Proportion Fatal (%)" = mean(fatal, na.rm = TRUE)*100, 
    "Proportion Injurious (%)" = mean(inj, na.rm = TRUE)*100, 
    "Proportion Harmless (%)"= mean(nohurt, na.rm = TRUE)*100,
  .groups = 'drop')
      
      
harm_selection <- c("Year", "Total Accidents")

if("fatal" %in% input$level_of_harm) {
  harm_selection <- c(harm_selection, "Fatal Accidents")
}
if("inj" %in% input$level_of_harm) {
  harm_selection <- c(harm_selection, "Injurious Accidents")
}
if("nohurt" %in% input$level_of_harm) {
  harm_selection <- c(harm_selection, "Harmless Accidents")
}

selected_data <- aggregated_data[, harm_selection, drop = FALSE]

selected_data
})
```

### Proportion of Accidents by Year & Harm

```{r}
renderTable({
  req(input$Year)
  
filtered_data <- dat %>% 
  filter(Year %in% input$Year) 
    
aggregated_data <- filtered_data %>%
  group_by(Year) %>%
  summarize(
    "Fatal Accidents" = sum(fatal, na.rm = TRUE),
    "Injurious Accidents" = sum(inj, na.rm = TRUE),
    "Harmless Accidents" = sum(nohurt, na.rm = TRUE),
    "Total Accidents" = sum(fatal, inj, nohurt, na.rm = TRUE),
    "Proportion Fatal (%)" = mean(fatal, na.rm = TRUE)*100, 
    "Proportion Injurious (%)" = mean(inj, na.rm = TRUE)*100, 
    "Proportion Harmless (%)"= mean(nohurt, na.rm = TRUE)*100,
  .groups = 'drop')
      
      
harm_selection <- c("Year") 

if("fatal" %in% input$level_of_harm) {
  harm_selection <- c(harm_selection, "Proportion Fatal (%)")
}
if("inj" %in% input$level_of_harm) {
  harm_selection <- c(harm_selection, "Proportion Injurious (%)")
}
if("nohurt" %in% input$level_of_harm) {
  harm_selection <- c(harm_selection, "Proportion Harmless (%)")
}

selected_data <- aggregated_data[, harm_selection, drop = FALSE]


colnames(selected_data) <- gsub("_", " ", colnames(selected_data))

selected_data
})
```

### Plotted Proportions of Accidents by Year & Harm

```{r}
renderPlot({
  req(input$Year)
  
filtered_data <- dat %>% 
  filter(Year %in% input$Year) 
    
aggregated_data <- filtered_data %>%
  group_by(Year) %>%
  summarize(
    "Fatal Accidents" = sum(fatal, na.rm = TRUE),
    "Injurious Accidents" = sum(inj, na.rm = TRUE),
    "Harmless Accidents" = sum(nohurt, na.rm = TRUE),
    "Total Accidents" = sum(fatal, inj, nohurt, na.rm = TRUE),
    "Proportion Fatal (%)" = mean(fatal, na.rm = TRUE)*100, 
    "Proportion Injurious (%)" = mean(inj, na.rm = TRUE)*100, 
    "Proportion Harmless (%)"= mean(nohurt, na.rm = TRUE)*100,
  .groups = 'drop')
      
      
harm_selection <- c("Year") 

if("fatal" %in% input$level_of_harm) {
  harm_selection <- c(harm_selection, "Proportion Fatal (%)")
}
if("inj" %in% input$level_of_harm) {
  harm_selection <- c(harm_selection, "Proportion Injurious (%)")
}
if("nohurt" %in% input$level_of_harm) {
  harm_selection <- c(harm_selection, "Proportion Harmless (%)")
}

selected_data <- aggregated_data[, harm_selection, drop = FALSE]


colnames(selected_data) <- gsub("_", " ", colnames(selected_data))

  long_data <- selected_data %>%
    tidyr::pivot_longer(-Year, names_to = "Category", values_to = "Proportion")
  
ggplot(long_data, aes(x = Year, y = Proportion, fill = Category)) +
  geom_area(position = 'stack') +
  labs(title = "Proportion of Harm Year-over-Year",
    x = "Year",
    y = "Proportion (%)",
  fill = "Level of Harm") +
  theme_minimal() 
})
```

Substance Use
=====================================  

Inputs {.sidebar}
-------------------------------------

```{r}
selectInput(
      inputId = "D1substanceFilter", 
      label = h3("Substance (Driver One):"),
      choices = c("All", "Alcohol", "Drugs", "Alcohol and Drugs", "No Apparent Influence"),
      selected = "All")


selectInput(
      inputId =  "timeOfDay", 
      label = h3("Select Time of Day:"),
      choices = c("All", "Morning Commute", "Evening Commute", "School Pickup", "Work", "Night", "Midnight to Dawn"),
      selected = "All")

```

Outputs
-------------------------------------

### Substance Use, Harm, Time Periods

```{r}
renderLeaflet({
  
  
  filtered_data <- dat %>%
    filter((input$D1substanceFilter == "All" | d1.substance == input$D1substanceFilter) &
           (input$timeOfDay == "All" | time.of.day == input$timeOfDay))
  
  col.vec <- ifelse(filtered_data$nohurt, 
                    "gray25", 
                    ifelse(filtered_data$inj, 
                           "orange", 
                           "red"))  
  
  point.size <- filtered_data$Totalinjuries + filtered_data$Totalfatalities
  
  crash.details <- paste0("Time: ", filtered_data$DateTime, "<br>",
                          "Total Fatalities: ", filtered_data$Totalfatalities, "<br>",
                          "Total Injuries: ", filtered_data$Totalinjuries, "<br>",
                          "Collision type: ", filtered_data$Collisionmanner)
  
  tempe <- leaflet(filtered_data) %>% 
    addProviderTiles("CartoDB.Positron") %>%
    setView(lng = -111.9278, lat = 33.4265, zoom = 13)
  
  tempe %>% 
    addCircles(lng = ~Longitude, 
               lat = ~Latitude,
               fillColor = ~col.vec, 
               fillOpacity = 0.5, 
               stroke = FALSE, 
               radius = ~50 * (1 + 0.33 * point.size),
               popup = ~crash.details) %>% 
    addLegend("topright",
              colors = c("red", "orange"),
              labels = c("Fatal", "Injurious"),
              title = "Accident Severity",
              opacity = 1)
  
})

``` 
About
===================================== 

Row 
-------------------------------------

### About this Dashboard

* The data utilized is available [here.](https://github.com/DS4PS/Data) For the most up-to-date data, visit the [City of Tempe](https://data.tempe.gov/datasets/tempegov::1-08-crash-data-report-detail/about) website. 
* This data contains traffic accident reports that occurred between 2012 and 2018 in Tempe, Arizona. The data is used by local government to reduce accidents resulting in fatalities and serious injury. It is used to improve safety in the city. 
* This dashboard serves as a tool to explore and examine traffic accidents in Tempe, Arizona. Diving into this dashboard will shed light on accident patterns by day, time, year, age, gender, type of transportation, and substance use. Although this is not a comprehensive view of all the data variables at play, it will serve its purpose for introductory exploration. Please note, "harmless" as it is used in the context of this dashboard, refers specifically to accidents that were not classifed as causing injuries or fatalities.  

Tab Breakdown:

* Day & Time: Use the map to visualize traffic accidents by the time of day and day of the week. 

* Age, Gender, & Method: This tab allows you to visually explore accidents based on the drivers characteristics, including age, gender, and type of transportation.  
* Comparisons: Create distinct maps for each driver to further isolate accidents using the variables of Age, Gender, & Method 
* Year & Harm: Compare rates of accidents across the years, including proportions of fatal, injurious, and harmless accidents.
* Substance Use: View a map of accidents by the involvement of substances within common traffic periods. 


### Dashboard Author

* [Sean West](https://www.linkedin.com/in/seanawest/) | Program Evaluation & Data Analytics, MS
* Recognition to Dr. Jesse Lecy for creating this dashboard from which additional tabs were built. 

Row
-------------------------------------

DATA DICTIONARY

```{r}

url.dd <- paste0("https://raw.githubusercontent.com",
                 "/DS4PS/cpp-526-fall-2019/master/l",
                 "abs/final-project/TempeTrafficAcc",
                 "identsDataDictionary.csv")

data.dictionary <- read.csv(url.dd, 
                            stringsAsFactors = FALSE)

data.dictionary$description <- stringi::stri_trans_general(data.dictionary$description,
                                                           "latin-ascii")

data.dictionary %>%
  select(column, description) %>%
  pander()

```

Data
=====================================  

```{r}

these.buttons <- c("copy", 
                   "csv", 
                   "pdf", 
                   "print")

renderDataTable({
  
  datatable(dat[1:100, ], 
            filter = "bottom", 
            rownames = FALSE, 
            fillContainer = TRUE, 
            style = "bootstrap",
            class = "table-condensed table-striped",
            extensions = "Buttons", 
            options = list(dom = "Bfrtip", 
                           buttons = these.buttons))

  })

```
